---
title: "ESP-195_Captstone_Project"
author: "Christian Seierup"
date: "2026-02-11"
output: html_document
indent: true
---

```{r setup, include=FALSE}
# Sets the default chunk to have "echo=FALSE"
# knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(fig.align = "center", warning = FALSE, message = FALSE)
```

```{r Load Libraries, include=FALSE}
# Add all libraries here.
# library(readxl)
# library(DT)
library(terra)
library(tidyr)
library(dplyr)
# library(fields) # For TPS interpolation
library(gstat) # For IDW Interpolation
```


## 1. Load Data

```{r 1. Load Data, eval=FALSE}
# 1.1 - GSPMD - GSP Monitoring Data
gwe_data_raw = read.csv("raw_data/gspmd/groundwater_level_data.csv")
gwe_sites_raw = read.csv("raw_data/gspmd/groundwater_level_sites.csv") 

# 1.2 - GAMA Data
gama_sites_raw = read.csv("raw_data/gama/gama_allwells.csv")
ddw_2000_09_raw = read.csv("raw_data/gama/ddw2000-2009.csv") # DDW samping data from 2000-2009
ddw_2010_19_raw = read.csv("raw_data/gama/ddw2010-2019.csv")
ddw_2020_pres_raw = read.csv("raw_data/gama/ddw2020-present.csv")

# 1.3 - Save Raw Data
save(gwe_data_raw, ddw_2000_09_raw, ddw_2010_19_raw, ddw_2020_pres_raw,
     file="data/my_Rdata/raw_data.Rdata")

save(gwe_sites_raw, gama_sites_raw,
     file="data/my_Rdata/raw_sites.Rdata")
```

## 2. Clean Data
### 2.1 GSPMD Data Cleaning
```{r, eval=FALSE}
load("data/my_Rdata/raw_data.Rdata")
load("data/my_Rdata/raw_sites.Rdata")
```

```{r, 2.1.1 Clean GSPMD Sites Data, eval=FALSE}
# 2.1.1 - GSPMD Sites
# GPSMD Groundwater Level Sample Sites
gwe_sites = gwe_sites_raw %>% 
  
  # Add Basin and Subbasin Information
  mutate(
    SUBBASIN_NUMBER = sapply(strsplit(BASIN_NAME,split=" ") , "[", 1), # unique subbasin number
    BASIN_NUMBER = substr(BASIN_NAME, 1, 5) # unique basin number
  ) %>% 
  
  # Select basins and Cols of interest
  select(SITE_CODE, SUBBASIN_NUMBER, BASIN_NUMBER, LATITUDE,
         LONGITUDE, WELL_DEPTH, LAST_GSE, SMC_MT) %>% # Columns of interest
  subset(BASIN_NUMBER %in% c("5-006", "5-021", "5-022")) %>%  # Only central valley basins
  
  # Tibble Format
  as_tibble()
```

```{r 2.1.2 Clean GSPMD Data, eval=FALSE}
# 2.1.2 - GSPMD Data
# GPSMD Groundwater Level Sample Data
gwe_data = gwe_data_raw %>%
  
  # Select Basins and Cols of interest
  subset(SITE_CODE %in% gwe_sites$SITE_CODE) %>%   # Only central valley wells
  select(SITE_CODE, MSMT_DATE, GWE) %>%  # Columns of interest
  
  # Tibble Format
  as_tibble()

# Add a post_SGMA column
gwe_data$post_SGMA = "Pre_SGMA"
gwe_data$post_SGMA[gwe_data$MSMT_DATE > "2014-01-01"] = "Post_SGMA"
```


### 2.2 GAMA/DDW Data Cleaning

```{r, 2.2.1 Clean GAMA Sites, eval=FALSE}
# 2.2.1 - GAMA Wells
# GAMA location information for all wells
gama_sites = gama_sites_raw %>% 
  
  # Add Basin and Subbasin Information
  mutate(
    SUBBASIN_NUMBER = stringr::str_extract(gm_gis_dwr_basin, "(?<=\\().*(?=\\))"),
    BASIN_NUMBER = substr(SUBBASIN_NUMBER, 1, 5) # "BASIN_NUMBER" contains the unique basin number
  ) %>% 
  
  # Select basins and Cols of interest
  subset(select=c("gm_well_id", "gm_latitude", "gm_longitude", "SUBBASIN_NUMBER", "BASIN_NUMBER")) %>% # Columns of interest
  subset(BASIN_NUMBER %in% c("5-006", "5-021", "5-022")) %>%  # Only central valley basins
  
  # Tibble Format
  as_tibble()
```

```{r, 2.2.2 Clean DDW Data, eval=FALSE}
# 2.2.2 - DDW Samples
# Columns and contaminants of interest
ddw_cols = c("gm_well_id", "gm_samp_collection_date", "gm_chemical_name", "gm_chemical_units", "gm_result", "gm_result_modifier", "gm_reporting_limit")
ddw_contaminants = c("Arsenic", "Chromium, Hexavalent (Cr6)", "Nitrate as N")

# Combine ddw datasets
ddw = rbind(ddw_2000_09_raw, ddw_2010_19_raw, ddw_2020_pres_raw) %>% 
  
  # Subset
  subset(gm_chemical_name %in% ddw_contaminants & # Contaminants of interest
         gm_well_id %in% gama_sites$gm_well_id, # Wells of interest
         select=ddw_cols) %>% # Columns of interest
  
  # Add Chemical Information
  mutate(
    gm_samp_collection_date = as.Date(gm_samp_collection_date), # Change Data type
    chemical = paste(gm_chemical_name, gm_chemical_units, sep="_") %>% 
      gsub(" ", "_", .) # Change all spaces to _ in chemical col
  ) %>% select(gm_well_id, gm_samp_collection_date, chemical, gm_result, 
               gm_result_modifier, gm_reporting_limit) %>% # Trim cols
  
  # Tibble Format
  as_tibble()

# Add a post_SGMA column
ddw$post_SGMA = "Pre_SGMA"
ddw$post_SGMA[ddw$gm_samp_collection_date > "2014-01-01"] = "Post_SGMA"
```

```{r, 2.3 Save Clean Data, eval=FALSE}
# 2.3 Save Data
save(gwe_data, ddw,
     file="data/my_Rdata/clean_data.Rdata")

save(gwe_sites, gama_sites,
     file="data/my_Rdata/clean_sites.Rdata")
```

## 3. Run Regression on Data
For each dataset, we want to determine the rate of change in each measured attribute (Groundwater Depletion, Pollutant Accumulation), at each well both before and after the implementation of SGMA to determine how SGMA affected these rates of change. 

### 3.1 GSPMD Regression Analysis
We will begin our analysis of the Groundwater Elevation data by dropping all NA
values, subsetting for only GWE measurements taken January 2000 and onwards,
and keeping only the relevant columns. We'll then group the measurements by 
whether they were taken Pre or Post SGMA, and by each unique well, and run a
simple linear regression on each group (this yields estimates of GWE change in
ft/day). We'll multiply each calculated slope by 365 to put it in units of
ft/year. 

```{r}
load("data/my_Rdata/clean_data.Rdata")
load("data/my_Rdata/clean_sites.Rdata")
```

```{r, 3.1.1 Subset GWE Data}
# 3.1.1 - Subset GWEData
gwe_data = gwe_data %>% 
  
  # Drop NA values
  drop_na(GWE) %>% 
  
  # Set date as a Date Data type
  mutate(
    MSMT_DATE = as.Date(MSMT_DATE)
  ) %>% 
  
  # Remove outlier well
  subset(
    MSMT_DATE >= "2000-01-01" & SITE_CODE != "357898N1192302W001", # Outlier
    select=c("SITE_CODE", "MSMT_DATE", "GWE", "post_SGMA")
  ) %>% 
  as_tibble()

# Result: gwe_data - a tibble of each GWE measurement
# 
# # A tibble: 243,342 × 4
#   SITE_CODE          MSMT_DATE    GWE post_SGMA
#   <chr>              <date>     <dbl> <chr>    
# 1 394559N1220854W001 2011-10-11  79.6 Pre_SGMA 
# 2 390174N1214630W003 2008-06-07  31.0 Pre_SGMA 
# 3 390284N1215155W001 2011-10-22  40.1 Pre_SGMA 
# 4 390174N1214630W002 2011-10-04  45.7 Pre_SGMA 
# 5 363500N1197800W003 2011-10-28 209.  Pre_SGMA 
# # ℹ 243,337 more rows
```

```{r, 3.1.2 GWE Regression}
# 3.1.2 - Run Regression, Select only slopes, trim columns
gwe_reg = gwe_data %>% 
  
  # Group by individual wells, and Before/After SGMA
  group_by(SITE_CODE, post_SGMA) %>% 
  
  # Run Regression for each group
  group_modify(~broom::tidy(lm(GWE ~ MSMT_DATE, data = .x)))%>% 
  
  # Select only the significant slopes
  subset(term=="MSMT_DATE" & p.value < .05) %>% 
  
  # Put slopes in units of change/year
  mutate(
    slope = estimate*365
  ) %>% 
  
  # Select only desired columns
  select(SITE_CODE, post_SGMA, slope) %>% 
  drop_na(slope)

# RESULT: gwe_reg - a tibble of pre/post SGMA rate of change in 
#                   groundwater elevation
# 
# # A tibble: 1,618 × 4
# # Groups:   SITE_CODE, post_SGMA [1,618]
#   SITE_CODE          post_SGMA   slope           p
#   <chr>              <chr>       <dbl>       <dbl>
# 1 349662N1189211W001 Post_SGMA -13.5   0.00485    
# 2 349836N1189228W001 Post_SGMA  -1.89  0.00335    
# 3 349976N1188463W001 Post_SGMA  -3.09  0.00000884 
# 4 350019N1188149W001 Post_SGMA   0.687 0.000000729
# 5 350033N1189366W001 Post_SGMA  -5.93  0.0313     
# # ℹ 1,613 more rows
```

```{r, 3.1.3 GWE Merge}
# Combine slopes with geodata
gwe_gdf = merge(gwe_reg, gwe_sites, by="SITE_CODE") %>% 
  
  # Change name of lat long 
  mutate(x = LONGITUDE, y = LATITUDE) %>% 
  
  # Select only desired columns
  select(SITE_CODE, post_SGMA, slope, x, y) %>% 
  as_tibble()

# RESULT: gwe_gdf - tibble of gwe_reg slopes with x and y coordinates attached
# 
# # A tibble: 1,618 × 6
#   SITE_CODE          post_SGMA   slope           p     x     y
#   <chr>              <chr>       <dbl>       <dbl> <dbl> <dbl>
# 1 349662N1189211W001 Post_SGMA -13.5   0.00485     -119.  35.0
# 2 349836N1189228W001 Post_SGMA  -1.89  0.00335     -119.  35.0
# 3 349976N1188463W001 Post_SGMA  -3.09  0.00000884  -119.  35.0
# 4 350019N1188149W001 Post_SGMA   0.687 0.000000729 -119.  35.0
# 5 350033N1189366W001 Post_SGMA  -5.93  0.0313      -119.  35.0
# # ℹ 1,613 more rows
```

### 3.2 DDW/GAMA Regression Analysis
We will now analyze the Division of Drinking water (DDW) and Groundwater
Ambient Monitoring and Assessment (GAMA) contaminant data. The analysis will be
nearly identical to our GSPMD analysis (3.1): we will drop all NA values, 
subsetting for only measurements taken January 2000 and onwards, and keeping
only the relevant columns. We'll then group the measurements by 
whether they were taken Pre or Post SGMA, and by each unique well, and run a
simple linear regression on each group (this yields estimates of GWE change in
ft/day). We'll multiply each calculated slope by 365 to put it in units of
ft/year. 

```{r, 3.2.1 Subset DDW data}
# 3.2.1 - Intro Subsetting
ddw_data = ddw %>% 
  
  # Drop NA Values
  drop_na(gm_result) %>% 
  
  # Set date as a Date Data type
  mutate(
    gm_date = as.Date(gm_samp_collection_date)
  ) %>% 
  
  # Select only desired columns
  subset(
    select=c("gm_well_id", 'gm_date', 'gm_result', 'chemical', "post_SGMA"),
    gm_samp_collection_date > "2000-01-01"
  ) %>% 
  as_tibble()

# Result: ddw_data - a tibble of each contaminant measurement
# 
# # A tibble: 1,047,850 × 5
#   gm_well_id        gm_date    gm_result chemical          post_SGMA
#   <chr>             <date>         <dbl> <chr>             <chr>    
# 1 CA0104010_003_003 2009-07-21       0.2 Nitrate_as_N_MG/L Pre_SGMA 
# 2 CA0110001_010_010 2008-07-23       1.3 Nitrate_as_N_MG/L Pre_SGMA 
# 3 CA0110001_019_019 2009-07-15       1.6 Nitrate_as_N_MG/L Pre_SGMA 
# 4 CA0110001_047_047 2008-07-16       1.5 Nitrate_as_N_MG/L Pre_SGMA 
# 5 CA0110003_014_014 2009-06-23       6.8 Nitrate_as_N_MG/L Pre_SGMA 
# # ℹ 1,047,845 more rows
```

```{r, 3.2.2 DDW Regression}
# 3.2.2 - Run Regression, Select only slopes, trim columns
ddw_reg = ddw_data %>% 
  
  # Group by individual wells, pre/post SGMA, and contaminant
  group_by(gm_well_id, post_SGMA, chemical) %>%
  
  # Run Regression for each group
  group_modify(~broom::tidy(lm(gm_result ~ gm_date, data = .x))) %>% 
  
  # Select only significant slopes
  subset(term=="gm_date" & p.value < .05) %>% 
  
  # Put slopes in units of change/year
  mutate(
    slope = estimate*365 #In units of x/year
  ) %>% 
  
  # Select only desired columns
  select(gm_well_id, post_SGMA, chemical, slope) %>% 
  drop_na(slope)

# RESULT: ddw_reg - a tibble of pre/post SGMA rate of change in 
#                   contaminant concentration
# 
# # A tibble: 5,008 × 5
# # Groups:   gm_well_id, post_SGMA, chemical [5,008]
#   gm_well_id        post_SGMA chemical            slope          p
#   <chr>             <chr>     <chr>               <dbl>      <dbl>
# 1 CA0105002_001_001 Post_SGMA Nitrate_as_N_MG/L -0.130  0.0327    
# 2 CA0105002_001_001 Pre_SGMA  Nitrate_as_N_MG/L -0.0916 0.00577   
# 3 CA0105016_001_001 Post_SGMA Nitrate_as_N_MG/L -0.0952 0.00000234
# 4 CA0300062_001_001 Pre_SGMA  Nitrate_as_N_MG/L -0.132  0.0489    
# 5 CA0300524_001_001 Post_SGMA Arsenic_UG/L      29.0    0.0328    
# # ℹ 5,003 more rows
```

```{r, 3.2.3 Merge DDW}
# Combine slopes with geodata
ddw_gdf = merge(ddw_reg, gama_sites, by="gm_well_id") %>% 
  
  # Change name of lat long
  mutate(x = gm_longitude, y = gm_latitude) %>% 
  
  # Select only desired Columns
  select(gm_well_id, post_SGMA, chemical, slope, x, y) %>% 
  as_tibble()

# RESULT: ddw_gdf - tibble of ddw_reg slopes with x and y coordinates attached
# 
# # A tibble: 4,739 × 7
#   gm_well_id        post_SGMA chemical             slope       p     x     y
#   <chr>             <chr>     <chr>                <dbl>   <dbl> <dbl> <dbl>
# 1 CA0105002_001_001 Post_SGMA Nitrate_as_N_MG/L -0.130   0.0327  -122.  37.8
# 2 CA0105002_001_001 Pre_SGMA  Nitrate_as_N_MG/L -0.0916  0.00577 -122.  37.8
# 3 CA0300062_001_001 Pre_SGMA  Nitrate_as_N_MG/L -0.132   0.0489  -121.  38.5
# 4 CA0300524_001_001 Post_SGMA Arsenic_UG/L      29.0     0.0328  -121.  38.3
# 5 CA0310008_002_002 Post_SGMA Nitrate_as_N_MG/L -0.00678 0.0160  -121.  38.2
# # ℹ 4,734 more rows
```

```{r, 3.3 Save Regression Geodataframes}
# 3.3 - Save Geodataframes
save(gwe_gdf, ddw_gdf,
     file="data/my_Rdata/gdfs.Rdata")
```


## 4. Interpolation (TPS and IDW)

```{r}
load("data/my_Rdata/gdfs.Rdata")
load("data/my_Rdata/clean_sites.Rdata")
```

### 4.0 Define functions, load Shapefiles

```{r, 4.0.2 IDW Function}
# 4.0 - Define IDW function
MyIDW = function(GDF, SHP, GSD, NAME){

# This function takes in a dataframe with columns 'x', 'y', and a 
# column of values ("slope") to perform a Inverse Distance Weighting (IDW) 
# regression. It also takes in a shapefile to get the bounds and
# CRS. It returns a raster of interpolated values with the bounds
# and CRS of the shapefile.
# 
# Inputs: 
#   GDF: dataframe with columns 'x' and 'y' for coordinates
#   SHP: SpatVector that sets the bounds for the interpolated raster
#   GSD: integer of desired Ground Sample Distance in units of the
#         SHP crs, probably Meters.
#   NAME: string of what to name the raster layer
#   
# Output:
#   my_idw: SpatRaster of IDW interpolated values with the 
#           same CRS of SHP, layer name of NAME
  
  
  # Project Shapefile
  SHP = SHP %>% 
    project("EPSG:3310") # NAD83 CA Teale Albers
  
  # Get Projected GDF Coordinates
  my_coords = as.matrix(GDF[,c("x", "y")]) %>%
    project("EPSG:4326", "EPSG:3310")

  # Add Projected coords back to GDF
  GDF$x = my_coords[,1]
  GDF$y = my_coords[,2]

  # Create Blank Raster to fill in
  my_rast = rast(ext(SHP) + 10000, crs = crs(SHP),
                 res=c(GSD, GSD)) # Ground Sample Distance
                                  # (GSD) in CRS units
  
  # Create gstat object w/ projected coordinates
  my_idw_gstat = gstat(id="slope_idw", formula=slope~1, locations=~x+y,
                            data=GDF, nmax=12, set=list(idp=2))
  
  # Run Interpolations
  my_idw = interpolate(my_rast, my_idw_gstat, debug.level=0, index=1)
  
  # Clip Interpolated rasters to basins area
  my_idw = crop(my_idw, SHP, mask=T)
  
  # Rename Raster
  names(my_idw) = NAME

  # Return IDW
  my_idw
}
```

```{r, 4.0.3 Open Shapefiles}
# 4.0.3 - Open Shapefiles
# Basins - Seperate
basins = vect("data/my_shapes/basins/i08_B118_CA_GroundwaterBasins.shp") %>% 
  
  # Only Central Valley Basins
  terra::subset(.$Basin_Numb %in% c("5-006", "5-021", "5-022"))

# Combined Basins + State Boundary (mostly for plotting insets)
ca_basins = rbind(
  vect("data/my_shapes/ca_state/CA_state.shp"), # First polygon is CA State Boundary
  aggregate(basins, dissolve=T) # Second polygon is the Entire CV basin
)

# Project Basins to the CRS of the well locations
basins = project(basins, "epsg:4326")
ca_basins = project(ca_basins, "epsg:4326") 
```

### 4.1 TPS for GWE

```{r, 4.1}
# 4.1.1 - Run TPS interpolation
# Pre SGMA TPS for GWE
gwe_pre_idw = MyIDW(GDF = gwe_gdf %>% subset(post_SGMA == "Pre_SGMA"),
                    SHP = ca_basins[2], GSD = 2000, NAME = "gwe_pre_ft_yr")

# Post SGMA TPS for GWE
gwe_post_idw = MyIDW(GDF = gwe_gdf %>% subset(post_SGMA == "Post_SGMA"),
                     SHP = ca_basins[2], GSD = 2000, NAME = "gwe_post_ft_yr")
```

### 4.2 IDW for DDW

```{r, 4.2.1}
# 4.2.1 - Run IDW for Nitrate
# Pre SGMA IDW for NO3
no3_pre_idw = MyIDW(GDF = ddw_gdf %>% subset(post_SGMA == "Pre_SGMA" & 
                                                chemical == "Nitrate_as_N_MG/L"),
                    SHP = ca_basins[2], GSD = 2000, NAME = "no3_pre_mg_L_yr")

# Post SGMA IDW for NO3
no3_post_idw = MyIDW(GDF = ddw_gdf %>% subset(post_SGMA == "Post_SGMA" &
                                                chemical == "Nitrate_as_N_MG/L"),
                     SHP = ca_basins[2], GSD = 2000, NAME = "no3_post_mg_L_yr")
```

```{r, 4.2.2}
# 4.2.2 - Run IDW for Hexavalent Chromium
# Pre SGMA IDW for Cr6
cr6_pre_idw = MyIDW(GDF = ddw_gdf %>% subset(post_SGMA == "Pre_SGMA" &
                                               chemical == "Chromium,_Hexavalent_(Cr6)_UG/L"),
                    SHP = ca_basins[2], GSD = 2000, NAME = "cr6_pre_ug_L_yr")

# Post SGMA IDW for Cr6
cr6_post_idw = MyIDW(GDF = ddw_gdf %>% subset(post_SGMA == "Post_SGMA" &
                                                chemical == "Chromium,_Hexavalent_(Cr6)_UG/L"),
                     SHP = ca_basins[2], GSD = 2000, NAME = "cr6_post_ug_L_yr")
```

```{r, 4.2.3}
# 4.2.3 - Run IDW for Arsenic
# Pre SGMA IDW for As
as_pre_idw = MyIDW(GDF = ddw_gdf %>% subset(post_SGMA == "Pre_SGMA" &
                                              chemical == "Arsenic_UG/L"),
                    SHP = ca_basins[2], GSD = 2000, NAME = "as_pre_ug_L_yr")

# Post SGMA IDW for As
as_post_idw = MyIDW(GDF = ddw_gdf %>% subset(post_SGMA == "Post_SGMA" &
                                               chemical == "Arsenic_UG/L"),
                     SHP = ca_basins[2], GSD = 2000, NAME = "as_post_ug_L_yr")
```


### 4.3 CV for GWE Data

Cross validation (CV) will be the same for all data. I'll create a function that automatically does the following:

Assign a group number (1-5) to all observation points

Loop through and create training and testing datasets (1 is testing, 2-5 are training)

Run the IDW with the training dataset

Retrieve the raster values at each point in the testing data

Calculate RMSE for that group

Repeat with a different group

```{r}
MyShuffleSplitCV = function(GDF, N){

  
# Set Seed/RMSE
  set.seed(101)
  my_errors = vector() # Blank vector to fill in later
  
# Assign a group number (1-5) to all observation points
  GDF$group = sample(rep(1:N, length.out=nrow(GDF)))

# Loop through and create training and testing datasets (1 is testing, 2-5 are training)
  for(test_group in 1:N){
    
  # Divide Training and Testing datasets
    train_groups = c(1:N)[-test_group]
  
    train = GDF %>% subset(group %in% train_groups)
    test = GDF %>% subset(group == test_group)
    
  # Run the IDW with the training dataset
    train_idw = MyIDW(GDF = GDF %>% subset(post_SGMA == "Pre_SGMA"),
                      SHP = ca_basins[2], GSD = 2000, NAME = "int_vals")
  
  # Retrieve the raster values at each point in the testing data
    
    # test SpatVector
    test_sv = vect(test, geom=c('x','y'), crs="EPSG:4326") %>% 
      project("EPSG:3310")
    
    # Values
    test = cbind(test, terra::extract(train_idw, test_sv)[2]) %>% as_tibble()
    
  
  # Calculate RMSE for that group
    errors = Metrics::mae(test$slope, test$int_vals)
    
  # Add RMSE to vector of RMSEs
    my_errors = my_erros %>% append(errors)
  }
  
  
# Return Average MAE
  return(mean(my_errors))
}
```

```{r}
jim = MyShuffleSplitCV(gwe_gdf, 10)
```


## 5. Zonal Statistics on Interpolations

```{r}
# load("data/my_Rdata/gdfs.Rdata")
# load("data/my_Rdata/clean_sites.Rdata")
```

### 5.1 Master Zonal Stats

```{r, 5.1 Zonal Stats}
# Reproject Basins
basins_zs = project(basins, "EPSG:3310")

# Combine all rasters
sgma_idw = rast(list(gwe_pre_idw, gwe_post_idw,
                     no3_pre_idw, no3_post_idw,
                     cr6_pre_idw, cr6_post_idw,
                     as_pre_idw, as_post_idw))

# Run Zonal Stats
sgma_zs = zonal(sgma_idw, basins_zs, fun="mean") %>%
  as_tibble() %>% 
  mutate(gwe_d_ft_yr = gwe_post_ft_yr - gwe_pre_ft_yr,
         no3_d_mg_L_yr = no3_post_mg_L_yr - no3_pre_mg_L_yr,
         cr6_d_ug_L_yr = cr6_post_ug_L_yr - cr6_pre_ug_L_yr,
         as_d_ug_L_yr = as_post_ug_L_yr - as_pre_ug_L_yr)

# Add values back to Shapefile
values(basins_zs) = values(basins_zs) %>% 
  
  # Columns of interest
  select(OBJECTID, Basin_Numb, Basin_Subb, Basin_Name, Basin_Su_1, Region_Off, Area_Acres) %>% 
  
  # Add Zonal Stats
  cbind(sgma_zs)
```

```{r, 5.2 Save Results}
# Save Zonal Stats File
write.csv(values(basins_zs), file="data/my_results/basins_zs.csv")
writeVector(basins_zs, file="data/my_results/basins_zs.gpkg", overwrite=T)
writeRaster(sgma_idw, file="data/my_results/sgma_idw.tiff", overwrite=T)
```


## 6. Plot Data

```{r}
# GWE
dev.new()
png(file="data/my_figures/GWE_interpolation_annual_change.png", width=2000, height=1250) # Initialize Image

# Initialize new plot
par(mfrow=c(1,3)) # Set arrangement of subplots
master_cex = 4 # For setting title and legend scale

# Set symbology/breaks
delta_col = colorRampPalette(c("#FF5D3D", "white", "#3DDBFF"))
gwe_col <- c('red', "#FF7070", "#FF8282", "#FFD9D9", 'white', "#B3DBFF", "#69B8FF", "#006FFF", "blue")  # create color ramp from red to blue
gwe_breaks = c(-20, -10, -5, -2, -.1, .1, 2, 5, 10, 26)

# Make Plot: Pre SGMA
plot(basins_zs, "gwe_pre_ft_yr",
     col=gwe_col, breaks=gwe_breaks, colNA="grey",
     main="Annual Trend in GWE Elevation\nPre SGMA",
     cex.main=master_cex,
     plg=list(x="bottomleft", title="dGWE\n(ft/year)", cex=master_cex))
terra::inset(ca_basins, loc="topright", scale=.5, col = c("grey", "white"))

# Add Plot: Post SGMA
plot(basins_zs, "gwe_post_ft_yr",
     col=gwe_col, breaks=gwe_breaks, colNA="grey",
     main="Annual Trend in GWE Elevation\nPost SGMA",
     cex.main=master_cex,
     plg=list(x="bottomleft", title="dGWE\n(ft/year)", cex=master_cex))
terra::inset(ca_basins, loc="topright", scale=.5, col = c("grey", "white"))

# Add Plot: Change pre vs post SGMA
plot(basins_zs, "gwe_d_ft_yr",
     col=delta_col(9), breaks=gwe_breaks, colNA="grey",
     main="Change in Annual Trend in GWE\nPre vs Post SGMA",
     cex.main=master_cex,
     plg=list(x="bottomleft", title="Change in dGWE\n(ft/year)", cex=master_cex))
terra::inset(ca_basins, loc="topright", scale=.5, col = c("grey", "white"))

dev.off() # Close file
```

```{r, 6.2.1}
# Nitrate
dev.new()
png(file="data/my_figures/NO3_Interpolation_annual_change.png", width=2000, height=1250) # Initialize Image

# Initialize new plot
par(mfrow=c(1,3)) # Set arrangement of subplots
master_cex = 4 # For setting title and legend scale

# Set symbology/breaks
delta_col = colorRampPalette(c("#FF5D3D", "white", "#3DDBFF"))
no3_col = colorRampPalette(c("chocolate", "white", "green3"))
no3_breaks = c(-.4, -.25, -.1, -.02, .02, .1, .25, .4)

# Make Plot: Pre SGMA
plot(basins_zs, "no3_pre_mg_L_yr",
     col=no3_col(7), breaks=no3_breaks, colNA="grey",
     main="Annual Trend in NO3 Concentration\nPre SGMA",
     cex.main=master_cex,
     plg=list(x="bottomleft", title="dNO3 Conc.\n(mg/L/year)", cex=master_cex))
terra::inset(ca_basins, loc="topright", scale=.5, col = c("grey", "white"))

# Add Plot: Post SGMA
plot(basins_zs, "no3_post_mg_L_yr",
     col=no3_col(7), breaks=no3_breaks, colNA="grey",
     main="Annual Trend in NO3 Concentration\nPost SGMA",
     cex.main=master_cex,
     plg=list(x="bottomleft", title="dNO3 Conc.\n(mg/L/year)", cex=master_cex))
terra::inset(ca_basins, loc="topright", scale=.5, col = c("grey", "white"))

# Add Plot: Change pre vs post SGMA
plot(basins_zs, "no3_d_mg_L_yr",
     col=delta_col(7), breaks=no3_breaks, colNA="grey",
     main="Change in Annual Trend in NO3 Conc.\nPre vs Post SGMA",
     cex.main=master_cex,
     plg=list(x="bottomleft", title="Change in\ndNO3 Conc.\n(mg/L/year)", cex=master_cex))
terra::inset(ca_basins, loc="topright", scale=.5, col = c("grey", "white"))

dev.off() # Close file
```

```{r, 6.2.2}
# 6.2.2 - Plot Chromium Results
dev.new()
png(file="data/my_figures/cr6_Interpolation_annual_change.png", width=2000, height=1250) # Initialize Image

# Initialize new plot
par(mfrow=c(1,3)) # Set arrangement of subplots

# Set symbology/breaks
cr6_col = colorRampPalette(c("chocolate", "#FFCF94", "white", "#91FCFF", "cyan"))
# cr6_breaks = c(-.4, -.3, -.2, -.1, .1, .2, .3, .4)
cr6_breaks = c(-1.2, -.8, -.4, -.1, .1, .4, .8, 1.2)

# Make Plot: Pre SGMA
plot(basins_zs, "cr6_pre_ug_L_yr",
     col=cr6_col(7), breaks=cr6_breaks, colNA="grey",
     main="Annual Trend in Cr6 Concentration\nPre SGMA",
     cex.main=master_cex,
     plg=list(x="bottomleft", title="dCr6 Conc.\n(mg/L/year)", cex=master_cex))
terra::inset(ca_basins, loc="topright", scale=.5, col = c("grey", "white"))

# Add Plot: Post SGMA
plot(basins_zs, "cr6_post_ug_L_yr",
     col=cr6_col(7), breaks=cr6_breaks, colNA="grey",
     main="Annual Trend in Cr6 Concentration\nPost SGMA",
     cex.main=master_cex,
     plg=list(x="bottomleft", title="dCr6 Conc.\n(mg/L/year)", cex=master_cex))
terra::inset(ca_basins, loc="topright", scale=.5, col = c("grey", "white"))

# Add Plot: Change pre vs post SGMA
plot(basins_zs, "cr6_d_ug_L_yr",
     col=delta_col(7), breaks=cr6_breaks, colNA="grey",
     main="Change in Annual Trend in Cr6 Conc.\nPre vs Post SGMA",
     cex.main=master_cex,
     plg=list(x="bottomleft", title="Change in\ndCr6 Conc.\n(mg/L/year)", cex=master_cex))
terra::inset(ca_basins, loc="topright", scale=.5, col = c("grey", "white"))

dev.off() # Close file
```

```{r, 6.2.3}
# 6.2.3 - Plot Arsenic Results
dev.new()
png(file="data/my_figures/AS_Interpolation_annual_change.png", width=2000, height=1250) # Initialize Image

# Initialize new plot
par(mfrow=c(1,3)) # Set arrangement of subplots

# Set symbology/breaks
as_col = colorRampPalette(c("chocolate", "#FFCF94", "white", "#C04FFF"))
delta_as_col = colorRampPalette(c("#FF5D3D", "#FF9985", "white", "#3DDBFF"))
as_breaks = c(-3.5, -2.5, -1.5, -.5, -.1, .1, .5, 1.5)

# Make Plot: Pre SGMA
plot(basins_zs, "as_pre_ug_L_yr",
     col=as_col(7), breaks=as_breaks, colNA="grey",
     main="Annual Trend in As Concentration\nPre SGMA",
     cex.main=master_cex,
     plg=list(x="bottomleft", title="dAs Conc.\n(mg/L/year)", cex=master_cex))
terra::inset(ca_basins, loc="topright", scale=.5, col = c("grey", "white"))

# Add Plot: Post SGMA
plot(basins_zs, "as_post_ug_L_yr",
     col=as_col(7), breaks=as_breaks, colNA="grey",
     main="Annual Trend in As Concentration\nPost SGMA",
     cex.main=master_cex,
     plg=list(x="bottomleft", title="dAs Conc.\n(mg/L/year)", cex=master_cex))
terra::inset(ca_basins, loc="topright", scale=.5, col = c("grey", "white"))

# Add Plot: Change pre vs post SGMA
plot(basins_zs, "as_d_ug_L_yr",
     col=delta_as_col(7), breaks=as_breaks, colNA="grey",
     main="Change in Annual Trend in As Conc.\nPre vs Post SGMA",
     cex.main=master_cex,
     plg=list(x="bottomleft", title="Change in\ndAs Conc.\n(mg/L/year)", cex=master_cex))
terra::inset(ca_basins, loc="topright", scale=.5, col = c("grey", "white"))

dev.off() # Close file
```














